---
title: "Changing_Lives_Bernard_Codebook"
author: "William Bernard"
date: "November 14, 2017"
output: pdf_document
---

# Study Design

The purpose of this study is to investigate measures of happiness and analyze the impact of loss/hardship on life satisfaction of individuals. This invesigation will also look at those who have suffered loss/hardship and analyze if a predisposition towards successfully navigating hardship effects overall life satisfaction. 

The sponsor of this study is the University of Washington 

Orignally, The ACL was sponsoreed by the National Institute of Aging (NIA), and the National Insitutes of Health (NIH). some later activities were supporteed by a "Robert Wood Johnson Foundation Investigators in Health Policy Award to James S. House" (ACL). 

The Data collection organization is named Americans' Changing Lives. 

This data set is a nationally representative cross-section survey of U.S. population aged 25+

This is a longitudinal study. Data was collected using a series of face-to-face or telephone interviews between 1986 - 2011. African Americans and people aged 60 + were over sampled at twice the rate of other groups. 

# Sampling

The population being sampled is the U.S. population aged 25 +.

A multistage stratified area probability sample with oversampling of Blacks and those aged 60 and older was used. 

# Variable Index

```{r, echo = FALSE}
library(dplyr)



load("C:\\Users\\William Bernard\\Desktop\\americans's_changing_lives_data_set\\ICPSR_04690\\DS0001\\04690-0001-Data.rda")

changing_lives_subset <- da04690.0001 %>% 
  select(V103, V104, V301, V314, V316, V317, V322, V323, V324, V328, V329, V330, V331, V332, V401, V416, V419, V420, V425, V429, V440, V441, V445, V446, V534, V613, V701, V706, V709, V710, V711, V732, V735, V934,V1002, V1006, V1007, V1009, V1010, V1022, V1501, V1504, V1505, V1508, V1520, V1524, V1601, V1602, V1603, V1605, V1609, V1619, V2024, V2027, V2029, V2033, V2200, V2308, V2309, V2310)

colnames(changing_lives_subset) <- c("sex_r", "age_r", "sat_life", "life_could_happier", "look_back_sat", "no_change_life", "sat_home", "sat_neighborhood", "safe_attack", "pos_self_att", "think_am_good", "feel_am_failure", "can_do_any", "pushed_around", "marital_stat", "one_mar_deal_death_spouse" ,"one_mar_deal_divorce", "times_widowed", "times_divorce", "two_plus_mar_deal_divorce", "child_die", "num_child_die", "child_death_unexp", "deal_child_death", "deal_death_parent", "deal_spouse_die", "ever_widowed", "death_unexp", "wid_sui_murd", "fatal_illness", "how_rec_death", "things_happen", "stronger_person", "see_psych", "felt_dep", "was_happy", "felt_lonely", "no_knows_well", "enjoyed_life", "event_cause_dep", "phys_attack", "deal_phys_attack", "life_threat_ill", "deal_ill", "lost_job_three_yr", "deal_job_loss", "accept_bad", "die_when_time", "part_god_plan", "bad_meant_be", "entitled_good_fut", "rew_afterlife", "widowed_y_n", "deal_death_sp_one_plus_mar", "ever_dev", "deal_dev", "life_sat_index", "child_die_three_yr", "assult_three_yr", "ill_threat_three_yr")

library(memisc)

changing_lives_subset_final <- as.data.set(changing_lives_subset)

```



```{r, echo = FALSE}

changing_lives_subset_final

changing_lives_subset_final <- within(changing_lives_subset_final, {
  description(sex_r) <- "Sex of Respondant"
  description(age_r) <- "Age of Respondant"
  description(sat_life) <- "Respondant satisfaction with life"
  description(life_could_happier) <- "Respondant thinks life could be happier"
  description(look_back_sat) <- "Looking back respondant is satisfied"
  description(no_change_life) <- "Respondant would not change their life"
  description(sat_home) <- "Respondant is satisfied with their home"
  description(sat_neighborhood) <- "Respondant is satisfied with their neighborhood"
  description(safe_attack) <- "Respondant feels safe from personal attack"
  description(pos_self_att) <-"Respondant has Positive self attitude "
  description(think_am_good) <- "Respondant thinks 'I am no good'"
  description(feel_am_failure) <- "Respondant feels 'I am a failure'"
  description(can_do_any) <- "Respondant thinks 'I can do anything'"
  description(pushed_around) <- "Respondant feels they are pushed around in life"
  description(marital_stat) <- "Marital Status of respondant"
  description(one_mar_deal_death_spouse) <- "One marriage, how did respondant deal with death of spouse"
  description(one_mar_deal_divorce) <- "One marriage, how did respondant deal with divorce "
  description(times_widowed) <- "Times respondant widowed" 
  description(times_divorce) <- "Number of respondant divorces"
  description(two_plus_mar_deal_divorce) <- "Two plus marriages, how did respondant deal with divorce"
  description(child_die) <- "Respondant ever have children die"
  description(num_child_die) <- "Number of respondant children died"
  description(child_death_unexp) <- "respondant's Child's death unexpected"
  description(deal_child_death) <- "How respondant dealt with child's death"
  description(deal_death_parent) <- "How respondant dealt with death of parent"
  description(deal_spouse_die) <- "Respondant deal well with death of spouse"
  description(ever_widowed) <- "Respondant ever widowed"
  description(death_unexp) <- "Was the death of the respondant's spouse unexpected?"
  description(wid_sui_murd) <- "Was the cause of the respondant's spouse's death suicide or murder?"
  description(fatal_illness) <- "how lond did respondant's spouse suffer from fatal illness?"
  description(how_rec_death) <- "How has the respondant recovered after the death of the spouse?"
  description(things_happen) <- "Respondant thinks things just happen"
  description(stronger_person) <- "Is the respondant stronger person now"
  description(see_psych) <- "Does the respondant see a psychiatrist"
  description(felt_dep) <- "Respondant was depressed"
  description(was_happy) <- "Respondant was happy"
  description(felt_lonely) <- "Respondant felt lonely"
  description(no_knows_well) <- "Respondant thinks 'no one knows me well'"
  description(enjoyed_life) <- "Respondant enjoys life"
  description(event_cause_dep) <- "An event caused respondants depression"
  description(phys_attack) <- "Respondant ever been physically attacked?"
  description(deal_phys_attack) <- "How did respondant deal with the phyisical attack"
  description(life_threat_ill) <- "Respondant ever have a lifethreatening illness"
  description(deal_ill) <- "How did the respondant deal with the lifethreatening illness"
  description(lost_job_three_yr) <- "Has the respondant lost a job in the last 3 years"
  description(deal_job_loss) <- "How did the responant react to losing the job in the last 3 years"
  description(accept_bad) <- "Does the respondant accept bad things?"
  description(die_when_time) <- "Does the respondant think you die when its time to die?"
  description(part_god_plan) <- "Does the respondant think everything is part of god's plan?"
  description(bad_meant_be) <- "Does the respondant think bad things are meant to be?"
  description(entitled_good_fut) <- "Does the respondant think they are entitled to a good future?"
  description(rew_afterlife) <- "Does the respondant think they will be rewarded in the afterlife?"
  description(widowed_y_n) <- "Has the respondant ever been widowed yes or no"
  description(deal_death_sp_one_plus_mar) <- "Respondants with one or more marriages, How did the respondant deal with the death of the spouse "
  description(ever_dev) <- "Has the respondant ever been divorced?"
  description(deal_dev) <- "Respondant divorced in last 3 years?"
  description(life_sat_index) <- "Life satisfaction index"
  description(child_die_three_yr) <- "Has the respondant had a child die in the last 3 years?"
  description(assult_three_yr) <- "Has the respondant been assulted in the last 3 years?"
  description(ill_threat_three_yr) <- "Has the respondant suffered from a life threatening illness in the last 3 years?"
  
})
```


```{r, echo = FALSE}
changing_lives_subset_final <- within(changing_lives_subset_final, {
measurement(sex_r) <- "nominal"
measurement(age_r) <- "ratio"
measurement(sat_life) <- "ordinal"
measurement(life_could_happier) <- "ordinal"
measurement(look_back_sat) <- "ordinal"
measurement(no_change_life) <- "ordinal"
measurement(sat_home) <- "ordinal"
measurement(sat_neighborhood) <- "ordinal"
measurement(safe_attack) <- "ordinal"
measurement(pos_self_att) <- "ordinal"
measurement(think_am_good) <- "ordinal"
measurement(feel_am_failure) <- "ordinal"
measurement(can_do_any) <- "ordinal"
measurement(pushed_around) <- "ordinal"
measurement(marital_stat) <- "ordinal"
measurement(one_mar_deal_death_spouse) <- "ordinal"
measurement(one_mar_deal_divorce) <- "ordinal"
measurement(times_widowed) <- "interval"
measurement(times_divorce) <- "interval"
measurement(two_plus_mar_deal_divorce) <- "ordinal"
measurement(child_die) <- "nominal"
measurement(num_child_die) <- "interval"
measurement(child_death_unexp) <- "nominal"
measurement(deal_child_death) <- "ordinal"
measurement(deal_death_parent) <- "ordinal"
measurement(deal_spouse_die) <- "ordinal"
measurement(ever_widowed) <- "nominal"
measurement(death_unexp) <- "nominal"
measurement(wid_sui_murd) <- "nominal"
measurement(fatal_illness) <- "nominal"
measurement(how_rec_death) <- "ordinal"
measurement(things_happen) <- "ordinal"
measurement(stronger_person) <- "ordinal"
measurement(see_psych) <- "nominal"
measurement(felt_dep) <- "ordinal"
measurement(was_happy) <- "ordinal"
measurement(felt_lonely) <- "ordinal"
measurement(no_knows_well) <- "ordinal"
measurement(enjoyed_life) <- "ordinal"
measurement(event_cause_dep) <- "nominal"
measurement(phys_attack) <- "nominal"
measurement(deal_phys_attack) <- "ordinal"
measurement(life_threat_ill) <- "nominal"
measurement(deal_ill) <- "ordinal"
measurement(lost_job_three_yr) <- "nominal"
measurement(deal_job_loss) <- "ordinal"
measurement(accept_bad) <- "ordinal"
measurement(die_when_time) <- "ordinal"
measurement(part_god_plan) <- "ordinal"
measurement(bad_meant_be) <- "ordinal"
measurement(entitled_good_fut) <- "ordinal"
measurement(rew_afterlife) <- "ordinal"
measurement(widowed_y_n) <- "nominal"
measurement(deal_death_sp_one_plus_mar) <- "ordinal"
measurement(ever_dev) <-"nominal"
measurement(deal_dev) <- "ordinal"
measurement(life_sat_index) <- "ratio"
measurement(child_die_three_yr) <-"nominal"
measurement(assult_three_yr) <- "nominal"
measurement(ill_threat_three_yr) <-"nominal"
})
```

```{r, echo = FALSE}
changing_lives_subset_final <- within(changing_lives_subset_final, {
labels(sex_r) <- c(
  "MALE"               = 1,
  "FEMALE"             = 2)
labels(sat_life) <- c(
  "COMPSAT"            = 1,
  "VERYSAT"            = 2,
  "SOMESAT"            = 3, 
  "NVERYSAT"           = 4,
  "NOTAASAT"           = 5)
labels(life_could_happier) <- c(
  "ST AGREE"           = 1,
  "AG SOME"            = 2,
  "DIS SOME"           = 3,
  "STR DIS"            = 4)
labels(look_back_sat) <- c(
  "ST AGREE"           = 1,
  "AG SOME"            = 2,
  "DIS SOME"           = 3,
  "STR DIS"            = 4)
labels(no_change_life) <- c(
  "ST AGREE"           = 1,
  "AG SOME"            = 2,
  "DIS SOME"           = 3,
  "STR DIS"            = 4)
labels(sat_home) <- c(
  "COMPSAT"            = 1,
  "VERYSAT"            = 2,
  "SOMESAT"            = 3, 
  "NVERYSAT"           = 4,
  "NOTAASAT"           = 5)
labels(sat_neighborhood) <- c(
  "COMPSAT"            = 1,
  "VERYSAT"            = 2,
  "SOMESAT"            = 3, 
  "NVERYSAT"           = 4,
  "NOTAASAT"           = 5)
labels(safe_attack) <- c(
  "VERYTRUE"           = 1,
  "MOSTTRUE"           = 2,
  "SOMETRUE"           = 3,
  "NOTTRUE"            = 4)
labels(pos_self_att) <- c(
  "ST AGREE"           = 1,
  "AG SOME"            = 2,
  "DIS SOME"           = 3,
  "STR DIS"            = 4)
labels(think_am_good) <- c(
  "ST AGREE"           = 1,
  "AG SOME"            = 2,
  "DIS SOME"           = 3,
  "STR DIS"            = 4)
labels(feel_am_failure) <- c(
  "ST AGREE"           = 1,
  "AG SOME"            = 2,
  "DIS SOME"           = 3,
  "STR DIS"            = 4)
labels(can_do_any) <- c(
  "ST AGREE"           = 1,
  "AG SOME"            = 2,
  "DIS SOME"           = 3,
  "STR DIS"            = 4)
labels(pushed_around) <- c(
  "ST AGREE"           = 1,
  "AG SOME"            = 2,
  "DIS SOME"           = 3,
  "STR DIS"            = 4)
labels(marital_stat) <- c(
  "MARRIED"            = 1,
  "SEPARATE"           = 2,
  "DIVORCED"           = 3,
  "WIDOWED"            = 4,
  "NEVMARRY"           = 5)
labels(one_mar_deal_death_spouse) <- c(
  "VERYWELL"           = 1,
  "QUITEWEL"           = 2,
  "SOMEWELL"           = 3,
  "NTTOWELL"           = 4)
labels(one_mar_deal_divorce) <- c(
  "VERYWELL"           = 1,
  "QUITEWEL"           = 2,
  "SOMEWELL"           = 3,
  "NTTOWELL"           = 4)
labels(two_plus_mar_deal_divorce) <- c(
  "VERYWELL"           = 1,
  "QUITEWEL"           = 2,
  "SOMEWELL"           = 3,
  "NTTOWELL"           = 4)
labels(child_die) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(child_death_unexp) <- c(
  "UNEXPECT"           = 1,
  "EXPECTED"           = 2)
labels(deal_child_death) <- c(
  "VERYWELL"           = 1,
  "QUITEWEL"           = 2,
  "SOMEWELL"           = 3,
  "NTTOWELL"           = 4)
labels(deal_death_parent) <- c(
  "VERYWELL"           = 1,
  "QUITEWEL"           = 2,
  "SOMEWELL"           = 3,
  "NTTOWELL"           = 4)
labels(deal_spouse_die) <- c(
  "STRAGREE"           = 1,
  "AGRSOME"            = 2,
  "DAGRSOME"           = 3,
  "ST DISAG"           = 4)
labels(ever_widowed) <- c(
  "EVRWIDOW"           = 1,
  "ALLOTHER"           = 2)
labels(death_unexp) <- c(
  "UNEXPECT"           = 1,
  "EXPECTED"           = 2,
  "UNDOCUMENTED CODE"  = 7)
labels(wid_sui_murd) <- c(
  "SUI/MURD"           = 1,
  "ALLOTHER"           = 2)
labels(fatal_illness) <- c(
  "AFEWDAYS"           = 1,
  "WEEKS"              = 2,
  "MONTHS"             = 3,
  "YRORMORE"           = 4,
  "UNDOCUMENTED CODE"  = 7)
labels(how_rec_death) <- c(
  "COMPLETE"           = 1,
  "ALM COMP"           = 2,
  "FAIR MCH"           = 3,
  "SOMEWHAT"           = 4,
  "NOTMUCH"            = 5)
labels(things_happen) <- c(
  "VERYTRUE"           = 1,
  "MOSTTRUE"           = 2,
  "SOMETRUE"           = 3,
  "NOTTRUE"            = 4)
labels(stronger_person) <- c(
  "VERYTRUE"           = 1,
  "MOSTTRUE"           = 2,
  "SOMETRUE"           = 3,
  "NOTTRUE"            = 4)
labels(see_psych) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(felt_dep) <- c(
  "HARDEVER"           = 1,
  "SOMETIME"           = 2,
  "MOSTTIME"           = 3)
labels(was_happy) <- c(
  "HARDEVER"           = 1,
  "SOMETIME"           = 2,
  "MOSTTIME"           = 3)
labels(felt_lonely) <- c(
  "HARDEVER"           = 1,
  "SOMETIME"           = 2,
  "MOSTTIME"           = 3)
labels(no_knows_well) <- c(
  "HARDEVER"           = 1,
  "SOMETIME"           = 2,
  "MOSTTIME"           = 3)
labels(enjoyed_life) <- c(
  "HARDEVER"           = 1,
  "SOMETIME"           = 2,
  "MOSTTIME"           = 3)
labels(event_cause_dep) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(phys_attack) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(deal_phys_attack) <- c(
  "VERYWELL"           = 1,
  "QUITEWEL"           = 2,
  "SOMEWELL"           = 3,
  "NTTOWELL"           = 4)
labels(life_threat_ill) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(deal_ill) <- c(
  "VERYWELL"           = 1,
  "QUITEWEL"           = 2,
  "SOMEWELL"           = 3,
  "NTTOWELL"           = 4)
labels(lost_job_three_yr) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(deal_job_loss) <- c(
  "VERYWELL"           = 1,
  "QUITEWEL"           = 2,
  "SOMEWELL"           = 3,
  "NTTOWELL"           = 4)
labels(accept_bad) <- c(
  "STRAGREE"           = 1,
  "AGRSOME"            = 2, 
  "DAGRSOME"           = 3, 
  "STRDAGR"            = 4)
labels(die_when_time) <- c(
  "STRAGREE"           = 1,
  "AGRSOME"            = 2, 
  "DAGRSOME"           = 3, 
  "STRDAGR"            = 4)
labels(part_god_plan) <- c(
  "STRAGREE"           = 1,
  "AGRSOME"            = 2, 
  "DAGRSOME"           = 3, 
  "STRDAGR"            = 4)
labels(bad_meant_be) <- c(
  "STRAGREE"           = 1,
  "AGRSOME"            = 2, 
  "DAGRSOME"           = 3, 
  "STRDAGR"            = 4)
labels(entitled_good_fut) <- c(
  "STRAGREE"           = 1,
  "AGRSOME"            = 2, 
  "DAGRSOME"           = 3, 
  "STRDAGR"            = 4)
labels(rew_afterlife) <- c(
  "STRAGREE"           = 1,
  "AGRSOME"            = 2, 
  "DAGRSOME"           = 3, 
  "STRDAGR"            = 4)
labels(widowed_y_n) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(deal_death_sp_one_plus_mar) <- c(
  "VERYWELL"           = 1,
  "QUITWELL"           = 2,
  "SOMEWHWE"           = 3,
  "NOT2WELL"           = 4)
labels(ever_dev) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(deal_dev) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(child_die_three_yr) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(assult_three_yr) <- c(
  "YES"                = 1,
  "NO"                 = 5)
labels(ill_threat_three_yr) <- c(
  "YES"                = 1,
  "NO"                 = 5)
})
```

```{r, echo = FALSE}

changing_lives_subset_final <- within(changing_lives_subset_final, {
missing.values(death_unexp) <- c(7)
missing.values(fatal_illness) <- c(7)
})
```

```{r, echo = FALSE}

codebook(changing_lives_subset_final)
```