---
title: "Lab 10 - Merging Data"
author: "William Bernard"
date: "November 2, 2017"
output:
  pdf_document: default
  html_document: default
---

Using your own dataset (which may include more than one table) carry out the following data cleaning steps. Knit together the PDF document and commit both the Lab 10 RMD file and the PDF document to Git. Push the changes to GitHub so both documents are visible in your public GitHub repository. 

1. For your poster project, do you have multiple tables you'd like to join together to create your complete dataset? If so, describe what each table represents. 

N/A

One table, Americans' changing lives represents the changing lives of americans. 


2. What is/are your primary key(s)? If you have more than one table in your data, what is/are your foreign key(s)? Do your primary key(s) and foreign key(s) have the same name? If not, what does this mean for the way you need to specify potential data merges?

the variables in my data are labeled uniformyly, V then number. Keys become less important since I will only be using one very large data set. 


3. If you do not need to merge tables to create your final dataset, create a new dataset from your original dataset with a `grouped_by()` summary of your choice. You will use this separate dataset to complete the following exercises. 

```{r}
library(dplyr)
library(tidyr)
library(tidyverse)

setwd("C:\\Users\\William Bernard\\Desktop\\William Bernard Poster Project v.2\\william_bernard_poster _project_v2")

# Read in your data with the appropriate function
changing_lives <- load("C:\\Users\\William Bernard\\Desktop\\americans's_changing_lives_data_set\\ICPSR_04690\\DS0001\\04690-0001-Data.rda")

  
data_subset <- da04690.0001 %>%
  select(V6, V7, V103, V104, V220, V221, V222, V223, V224, V225, V301, V302, V303, V304, V305, V306, V307, V308, V309, V310, V311, V314, V315, V316, V317, V322, V323, V325, V326, V328, V329, V330, V331, V332, V333, V334, V335, V401, V402, V405, V406, V407, V408, V410, V416, V419, V420, V425, V430, V431, V432, V433, V434, V437, V438, V440, V441, V445, V446)

group_by(data_subset, V301)
```

If you are merging separate tables as part of your data manipulation process, are your keys of the same data type? If not, what are the differences? Figure out the appropriate coercion process(es) and carry out the steps below. 

The keys are the same. 


4. Perform each version of the mutating joins (don't forget to specify the `by` argument) and print the results to the console. Describe what each join did to your datasets and what the resulting data table looks like. For those joining two separate datasets, did any of these joins result in your desired final dataset? Why or why not?
```{r}
data_table_1 <- data_subset %>%
  select(V301, V305, V311, V432, V408)

data_table_2 <- data_subset %>%
  select(V301, V420, V335, V309, V437)

as_tibble(left_join(data_table_1, data_table_2, by = "V301"))

as_tibble(right_join(data_table_1, data_table_2, by = "V301"))

as_tibble(inner_join(data_table_1, data_table_2, by = "V301"))

as_tibble(full_join(data_table_1, data_table_2))
```

5. Do the same thing with the filtering joins. What was the result? Give an example of a case in which a `semi_join()` or an `anti_join()` might be used with your primary dataset

```{r}
as_tibble(semi_join(data_table_1, data_table_2, by = "V301"))

```

I would most likely not use this function, but If I was, it would be to check if there are similar rows in another data frame. 

```{r}

as_tibble(anti_join(data_table_1, data_table_2, by = "V301"))

```

I would use this to see how useful a potential data set would be to me. 


6. What happens when you apply the set operations joins to your tables? Are these functions useful for you for this project? Explain why or why not. If not, give an example in which one of them might be usefully applied to your data. 

```{r}
as_tibble(intersect(data_table_1, data_table_1))

as_tibble(union(data_table_1, data_table_1))

as_tibble(setdiff(data_table_1, data_table_1))
```

none of these functions are particularly useful for this project, I have one data set. If I got my hands on other data sets, union could be useful when looking at the variables of other data sets. Quickly finding the overlap with this function would be very helpful. 

7. If you have any reason to compare tables, apply `setequal()` below. What were the results? 

N/A


8. What is the purpose of binding data and why might you need to take extra precaution when carrying out this specific form of data merging? If your data requires any binding, carry out the steps below and describe what was accomplished by your merge.


binding basically pastes data sets on to one another. you could mess with the total observations of some of the variables if the data doesnt line up right, which would severly hinder your work. 

N/A

9. Do you need to merge multiple tables together using the same type of merge? If so, utilize the `reduce()` function from the `purrr` package to carry out the appropriate merge below. 


N/A


10. Are there any other steps you need to carry out to further clean, transform, or merge your data into one, final, tidy dataset? If so, describe what they are and carry them out below.

N/A












